### 2022-01-10 cleave.tex
trimming

In this section, we present and detail our \ac{CLEAVE} framework for the performance evaluation of \acfp{NCS}, with a particular focus on edge deployments.
The framework is distributed as \emph{free and open-source} software, and can be found online in the \emph{GitHub} organization of the \emph{KTH ExPECA} research group\footnote{\url{https://github.com/KTH-EXPECA/CLEAVE}}.
Our design follows a virtualized emulation approach; we provide a \emph{framework} for the real-time \emph{emulation} of physical control systems which interact with \emph{real} networks.
This allows for great flexibility when it comes to the components of the control system, as these can easily be switched out in the fully emulated approach, while maintaining the realism of arguably the most complex and limiting component of and edge \acp{NCS}, namely the network itself.
In other words, \ac{CLEAVE} is a platform with an easy-to-use \ac{API} on which a multitude of \ac{NCS} types can be emulated.
Our vision is to furthermore provide an open library of \ac{NCS} implementations in the future, this way making the tool even more accessible to the community.

@@

\begin{figure*}
    \centering
    \includegraphics[width=.8\textwidth]{CLEAVE_NCS_structure}
    \caption{
        Structure of an emulated \ac{NCS} on top of CLEAVE.\@
        The State object implements the discrete-time evolution of the Plant.
        Actuated Variables are updated from the values obtained from the Actuator objects at the beginning of each time step and are used to update the State before each time-step.
        Sensed Variables are sampled at the end of each time-step and passed on to the Sensor objects, which process them (e.g. to add noise) and send the processed values to the Controller.
        Controller calculates actuation values and sends these to the Actuators.
        Finally, the Actuators process the actuation values and hold the results until they are read at the next time-step.
    }\label{fig:cleave:ncs:struct}
\end{figure*}

@@

The first step in developing an \ac{NCS} emulation on top of \ac{CLEAVE} is the definition of a Plant which represents the physical system to be controlled.

In abstract terms, Plants in \ac{CLEAVE} can be understood to be composed of
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item\label{item:plant:state} an emulation of a physical system which is updated in a discrete-time fashion
    \item\label{item:plant:actuators} procedures to modify or act upon the physical emulation at each time-step
    \item\label{item:plant:sensors} procedures to sample relevant values from the physical emulation at each time-step
    \item\label{item:plant:network} procedures to interact with the network and send and receive data from the Controller.
\end{enumerate*}

While \cref{item:plant:network} is handled transparently by CLEAVE, by hiding it through layers of abstraction, the rest of the items need to be provided in more explicit ways by the user.

For \cref{item:plant:state}, the user needs to extend an abstract base class \texttt{State} provided by the framework.
This class provides the base functionality and defines a simple API for the implementation of discrete-time emulations by
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item automatically tracking variables that can be modified and/or sampled by the framework
    \item defining a single abstract method \texttt{State.advance()} which is invoked by the framework on each time step of the emulation.
\end{enumerate*}

\Cref{item:plant:actuators,item:plant:sensors} are in the same way provided by the user through the extension of the \texttt{Actuator} and \texttt{Sensor} abstract base classes, respectively.
\texttt{Actuator} objects process the outputs of the Controller and update the values of the \texttt{State} accordingly.
Conversely, \texttt{Sensor} objects sample relevant values from the \texttt{State} object, process them, and finally pass them on to the Controller through the network.

@@

After the Plant has been defined, users need to implement a Controller for it.
This is once again done by extending an abstract base class, \texttt{Controller}, provided by CLEAVE.\@
This class defines a single abstract method, \texttt{Controller.process()} which is called by the framework whenever samples are provided by the Plant.
When invoked, it is passed a mapping of property names to sampled values.
Implementations of this method should return yet another mapping of actuated property names to target values.
Users are otherwise free to implement any kind of logic or process within extending classes.

@@

Once the core components for the emulation have been defined and implemented, the emulation needs to be configured to register them with the framework.
This is done through fully-featured Python scripts which simply define a number of required (and some optional) top-level variables.
An example of such a configuration file can be seen in \cref{lst:config:plant}.
These Python configuration files are then passed as arguments to the \texttt{cleave} bash command provided by the framework, which sets up and executes the emulation.
The use of Python scripts for configuration allows for a high degree of flexibility and complexity in the definition of emulation setups.

@@

\begin{figure}
    \centering
    \includegraphics[width=.7\columnwidth]{images/inverted_pendulum.png}
    \caption{
        The two-dimensional inverted pendulum system.
        The objective of the system is to balance the vertically free-swinging pendulum by moving the cart.
    }\label{fig:invpend}
\end{figure}

@@

\begin{listing}
    \begin{minted}[linenos=true,
        breaklines, 
        breakafter=d,
        fontsize=\small]{Python}
from cleave.api.plant import SimpleConstantActuator, SimpleSensor
from cleave.impl.inverted_pendulum import InvPendulumState

>>>>

In this section, we present and detail our \ac{CLEAVE} framework for the performance evaluation of \acfp{NCS}, with a particular focus on edge deployments.
The framework is distributed as \emph{free and open-source} software, and can be found online in the \emph{GitHub} organization of the \emph{KTH ExPECA} research group\footnote{\url{https://github.com/KTH-EXPECA/CLEAVE}}.
Our design follows a virtualized emulation approach; we provide a framework for the real-time \emph{emulation} of physical control systems which interact with \emph{real} networks.
This allows for great flexibility when it comes to the components of the control system, as these can easily be switched out in the fully emulated approach, while maintaining the realism of the network, arguably the most complex and limiting component of and edge \acp{NCS}.
In other words, \ac{CLEAVE} is a platform with an easy-to-use \ac{API} on which a multitude of \ac{NCS} types can be emulated.
Our vision is furthermore to provide an open library of \ac{NCS} implementations in the future, this way making the tool even more accessible to the community.

@@

\begin{figure*}
    \centering
    \includegraphics[width=.8\textwidth]{CLEAVE_NCS_structure}
    \caption{
        Structure of an emulated \ac{NCS} in \ac{CLEAVE}.\@
        The State object implements the discrete-time evolution of the Plant.
        Actuated Variables are updated from the values obtained from the Actuator objects at the beginning of each time step and are used to update the State before each time-step.
        Sensed Variables are sampled at the end of each time-step and passed on to the Sensor objects, which process them (e.g. to add noise) and send the processed values to the Controller.
        Controller calculates actuation values and sends these to the Actuators.
        Finally, the Actuators process the actuation values and hold the results until they are read at the next time-step.
    }\label{fig:cleave:ncs:struct}
\end{figure*}

@@

The first step in developing an \ac{NCS} emulation on top of \ac{CLEAVE} is the definition of a Plant which represents the physical system to be controlled.
In abstract terms, Plants in \ac{CLEAVE} can be understood to be composed of
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item\label{item:plant:state} an emulation of a physical system which is updated in a discrete-time fashion
    \item\label{item:plant:actuators} procedures to modify or act upon the physical emulation at each time-step
    \item\label{item:plant:sensors} procedures to sample relevant values from the physical emulation at each time-step
    \item\label{item:plant:network} procedures to interact with the network and send and receive data from the Controller.
\end{enumerate*}

While \cref{item:plant:network} is handled transparently by CLEAVE, by hiding it through layers of abstraction, the rest of the items need to be provided in more explicit ways by the user.
For \cref{item:plant:state}, the user needs to extend an abstract base class \texttt{State} provided by the framework.
This class provides the base functionality and defines a simple API for the implementation of discrete-time emulations by
\begin{enumerate*}[itemjoin={{; }}, itemjoin*={{; and }}]
    \item automatically tracking variables that can be modified and/or sampled by the framework
    \item defining a single abstract method \texttt{State.advance()} which is invoked by the framework on each time step of the emulation.
\end{enumerate*}
\cref{item:plant:actuators,item:plant:sensors} are in the same way provided by the user through the extension of the \texttt{Actuator} and \texttt{Sensor} abstract base classes, respectively.
\texttt{Actuator} objects process the outputs of the Controller and update the values of the \texttt{State} accordingly.
Conversely, \texttt{Sensor} objects sample relevant values from the \texttt{State} object, process them, and finally pass them on to the Controller through the network.

@@

After the Plant has been defined, a Controller needs to be implemented for it.
This is once again done by extending an abstract base class, \texttt{Controller}, provided by CLEAVE.\@
This class defines a single abstract method, \texttt{Controller.process()} which is called by the framework whenever samples are provided by the Plant.
When invoked, it is passed a mapping of property names to sampled values.
Implementations of this method should return yet another mapping of actuated property names to target values.
Users are otherwise free to implement any kind of logic or process within extending classes.

@@

Once the core components for the emulation have been defined and implemented, the emulation needs to be configured to register them with the framework.
This is done through fully-featured Python scripts which simply define a number of required (and some optional) top-level variables.
An example of such a configuration file can be seen in \cref{lst:config:plant}.
These Python configuration files are then passed as arguments to the \texttt{cleave} command line provided by the framework, which sets up and executes the emulation.
The use of Python scripts for configuration allows for a high degree of flexibility and complexity in the definition of emulation setups.

@@

\begin{figure}
    \centering
    \includegraphics[width=.7\columnwidth]{images/inverted_pendulum.png}
    \caption{
        The 2d inverted pendulum system, the objective of which is to balance the vertically free-swinging pendulum on the cart.
    }\label{fig:invpend}
\end{figure}

@@

\begin{listing}
    \begin{minted}[linenos=false,
        breaklines, 
        breakafter=d,
        fontsize=\footnotesize]{Python}
from cleave.api.plant import SimpleConstantActuator, SimpleSensor
from cleave.impl.inverted_pendulum import InvPendulumState

### 2022-01-10 experiments.tex
trimming

We depict our experimental edge setup in \cref{fig:cleave:expsetup}.
It consists of \num{10} Raspberry Pi 4B  clients connected wirelessly to an IEEE 802.11n \ac{AP}; connected to the Ethernet backbone of this \ac{AP} is a general-purpose Cloudlet.
On top of this physical architecture we configure a Docker Swarm\footnote{Swarm mode overview: \url{https://docs.docker.com/engine/swarm/}} managed centrally from the Cloudlet.
For each experimental scenario we deploy a number of control loops inside this Swarm, executing plant containers exclusively in the Raspberry Pi clients (up to a single plant per client) and controller containers in the Cloudlet.
Plants and controllers communicate using the \ac{UDP} over an overlay network which sits on top of and abstracts away the real network configuration.
Additionally, to obtain baseline results without the stochastic effects of the network, we employ a secondary ``local-only'' setup, in which plants and controllers are executed co-located on the Cloudlet.

@@

As expected, higher sampling rates directly correlate to better quality of control (with the exception of the \SIlist{40;60}{\hertz}, \SI{50}{\milli\second} scenarios --- we attribute this to randomness and the relatively short duration of the experiments).
As sampling rates increased, the system was able to reach stability at higher \ac{RTT} latencies.
These initial results already hint at interesting consequences for \ac{NCS} deployments on the edge.
For instance, it is clear from \cref{fig:single:stability} that network delays can, to a certain extent, be compensated for by increasing the sampling rate of the system.
A corollary of this is that, conversely, at lower network latencies \acp{NCS} are able to stabilize at lower sampling rates.
Adaptive sampling might thus be a straightforward method for optimizing \ac{NCS} resource utilization and energy consumption at the edge.

@@

These results stem from the contention for network resources, and hints at important trade-offs system designers will have to take into consideration when designing and developing \acl{NCS} for deployment on the edge.
The edge will be \emph{multi-tenant} and \emph{multi-instance}, and thus any edge \ac{NCS} deployment will have to be designed with shared resources in mind.
\acp{NCS} will have to dynamically adapt to system limits in order to maintain quality of control.

>>>>

We detail our experimental edge setup in \cref{fig:cleave:expsetup,tab:hardware}.
It consists of \num{10} Raspberry Pi 4B  clients connected wirelessly to an IEEE 802.11n \ac{AP}; connected to the Ethernet backbone of this \ac{AP} is a general-purpose Cloudlet.
On top of this physical architecture we configure a Docker Swarm\footnote{Swarm mode overview: \url{https://docs.docker.com/engine/swarm/}} managed centrally from the Cloudlet.
For each experimental scenario we deploy a number of control loops inside this Swarm, executing plant containers exclusively in the Raspberry Pi clients (up to a single plant per client) and controller containers in the Cloudlet.
Plants and controllers communicate using the \ac{UDP} over an overlay network which sits on top of and abstracts away the real network configuration.
Additionally, to obtain baseline results without the stochastic effects of the network, we employ a secondary ``local-only'' setup, in which plants and controllers are executed co-located on the Cloudlet.

@@

As expected, higher sampling rates directly correlate to better quality of control (with the exception of the \SIlist{40;60}{\hertz}, \SI{50}{\milli\second} scenarios --- we attribute this to randomness and the relatively short duration of the experiments).
As sampling rates increased, the system was able to reach stability at higher \acp{RTT}.
These initial results already hint at interesting consequences for \ac{NCS} deployments on the edge.
For instance, it is clear from \cref{fig:single:stability} that network delays can, to a certain extent, be compensated for by increasing the sampling rate of the system.
A corollary of this is that, conversely, at lower network latencies \acp{NCS} are able to stabilize at lower sampling rates.
Adaptive sampling might thus be a straightforward method for optimizing \ac{NCS} resource utilization and energy consumption at the edge.

@@

These results stem from the contention for network resources, and hints at important trade-offs system designers will have to take into consideration when designing and developing \acp{NCS} for deployment on the edge.
The edge will be \emph{multi-tenant} and \emph{multi-instance}. 
\acp{NCS} deployments will have to be designed with shared resources in mind, and given the complexity of these systems, experimental tools like \ac{CLEAVE} will be key for their succesful adoption and massification.

### 2022-01-10 intro.tex
trimming

\textcite{Zoppi2020NCSBench} made the first (and to the best of our knowledge, the only) attempt at tackling this challenge with their \emph{NCSbench} platform.
\emph{NCSbench} is an open-source \ac{NCS} benchmarking platform, built on top of the LEGO\textregistered{}\ Mindstorms EV3 Core Set\texttrademark{}\ platform, with stated goals of reproducibility and ease of use.
It achives this thanks to its low-cost, accessibility, and ease of reconfiguration.

@@

In this paper, we present the first fully-software-based \ac{NCS} benchmarking framework with scalability and repeatability as its main focus, built with the Edge in mind.
In the years since the advent of the Edge computing paradigm, more and more variations of it have appeared in literature as the paradigm starts to be adopted by industry.
``Near'', ``far'', ``core'', and ``telco'' edge, among others, are all terms which describe variations of the original concept and which are becoming ubiquitous in new research.
While the core idea of Edge computing is widely accepted as fundamental for pervasive \acp{NCS} in general, understanding the strenghts, and weaknesses of such different edge concepts is of paramount importance.

Our framework, which we name \ac{CLEAVE}, aims to simplify the repeatable and scalable benchmarking of such systems. 
It is built on top of the \emph{Python 3.8} language, making it highly extensible and able to harness the hundreds-of-thousands of already existing user-provided libraries and packages.
Additionally, it is fully compatible with \emph{containerization} technologies such as \emph{Docker}\footnote{Docker Engine: \url{https://www.docker.com/}}, making it suitable for automated deployment and benchmarking on industry-standard cloud and (more importantly) edge setups using container orchestration technologies.

The rest of this paper is structured as follows.
\Cref{sec:approach} presents the design principles of the framework.
In \cref{sec:experiments}, we present a series of experiments which validate the utility, flexibility, and repeatability of \ac{CLEAVE}.
Finally,in \cref{sec:conclusion} we conclude this paper and briefly discuss future work.

>>>>

\textcite{Zoppi2020NCSBench} made the first (and to the best of our knowledge, the only) attempt at tackling this challenge with their \emph{NCSbench} platform.
\emph{NCSbench} is an open-source \ac{NCS} benchmarking platform, built on top of the LEGO\textregistered{}\ Mindstorms EV3 Core Set\texttrademark{}\ platform, with stated goals of reproducibility and ease of use.
It achieves this thanks to its low-cost, accessibility, and ease of reconfiguration.

@@

In this paper, we present the first fully-software-based framework for scalable and repeatable benchmarking of edge-native \ac{NCS}.
As the Edge computing paradigm starts to be adopted by industry, more and more variations of it have started to appear in literature.
``Near'', ``far'', ``core'', and ``telco'' edge, among others, are all terms which describe variations of the original concept and which are becoming ubiquitous in new research.
While the core idea of Edge computing is widely accepted as fundamental for pervasive \acp{NCS} in general, understanding the strenghts and weaknesses of such different edge concepts is of paramount importance.

Our framework, which we name \ac{CLEAVE}, aims to simplify the repeatable and scalable benchmarking of such systems. 
It is built on top of the \emph{Python 3.8} language, making it highly extensible and able to harness the hundreds-of-thousands of already existing user-provided libraries and packages.
Additionally, it is compatible with \emph{containerization} technologies such as \emph{Docker}\footnote{Docker Engine: \url{https://www.docker.com/}}, making it suitable for automated deployment and benchmarking on industry-standard cloud and (more importantly) edge setups using container orchestration technologies.

The rest of this paper is structured as follows.
\cref{sec:approach} presents the design principles of the framework.
In \cref{sec:experiments}, we present a series of experiments which validate the utility, flexibility, and repeatability of \ac{CLEAVE}.
Finally, in \cref{sec:conclusion} we conclude this paper and briefly discuss future work.
